---
title: "RWorkshopJune16"
author: "AES"
date: '2022-06-15'
output: html_document
---

## This is our document for the R Workshop on Exploratory Data Analysis.


### Before you can use a library (aka package of functions), you have to initially install the library/package
```{r echo=FALSE}
#install.packages("dplyr")
#install.packages("corrplot")
#install.packages("GGally")
#install.packages("tidycensus")
#install.packages("mapview)
```


### Load libraries
```{r warning=FALSE,message=FALSE}
library(dplyr)
library(corrplot)
library(GGally)
library(tidycensus)
library(mapview)
```


### Load mtcars (dataset we used last week with Joe Risi)
```{r}
data("mtcars")
```


### Basic info about the dataset
```{r}
nrow(mtcars) # number of rows
ncol(mtcars) # number of columns
head(mtcars) # first five rows
tail(mtcars) # last five rows
```


### if you don't remember what a function is or you're stumped on how to use it:
```{r}
#?nrow
```


### summary() & associated functions
summaries of data trends
```{r}
# initial look at our data
summary(mtcars)

# look at just one column of data
summary(mtcars$mpg)

# look at just one function of one column of data
min(mtcars$mpg)
max(mtcars$mpg)
median(mtcars$mpg)
mean(mtcars$mpg)
sd(mtcars$mpg)
```


### table() & prop.table()
engine types, gears, and types vs. gears
```{r}
table(mtcars$vs) # 14 with a straight engine
table(mtcars$gear) # 15 with 3 gears, 12 with 4, 5 with 5
table(mtcars$vs,mtcars$gear)

# use prop.table() to see proportions
prop.table(table(mtcars$vs))
prop.table(table(mtcars$gear))
prop.table(table(mtcars$vs,mtcars$gear))

# prop.table requires a table as its input e.g., "table(mtcars$vs)"
# instead of typing it all out, we could save this table as an object
vs_tab <- table(mtcars$vs)
vs_tab
prop.table(vs_tab)

# what are the tables below telling us?
prop.table(table(mtcars$vs,mtcars$gear),1)
prop.table(table(mtcars$vs,mtcars$gear),2)
```


### aggregate() & group_by()
```{r}
# let's say we want to know all the means of the data by engine type
aggregate(mtcars,by=list(mtcars$vs),FUN=mean)

# how do the gear groups differ in mpg?
aggregate(mtcars$mpg,by=list(mtcars$gear),FUN=mean)

# there are multiple ways to achieve the same thing! here's a tidyverse example:
mtcars %>% group_by(gear) %>% summarise_at(vars(mpg),mean)
```


### hist() & boxplot() & plot()
simple graphical representations of data
```{r}
hist(mtcars$mpg)

boxplot(mpg ~ vs,mtcars)
boxplot(mpg ~ gear,mtcars)

# sometimes Rmd doesn't like adding stuff to plots. you'll have to run the next two lines manually in the console:
plot(mpg ~ wt,mtcars)
abline(lm(mpg ~ wt,mtcars))
# if you "knit" this document, this code will work fine!
```
Megan and Anna will show how to make prettier and more useful plots in July!


### cor() & corrplot() & ggpairs()
```{r}
cor(mtcars)
#?cor # uses pearson method by default

corrplot(cor(mtcars))

# what types of options can we add to corrplot?
corrplot(cor(mtcars),method="number")
corrplot(cor(mtcars),method="color")
corrplot(cor(mtcars),method="color",type="upper")

# or maybe we only care about a couple variables:
corrplot(cor(mtcars %>% select(mpg,vs,gear,wt)),method="color",type="upper")

# ggpairs is a fancy function that combines a lot of the functions we talked about today:
ggpairs(mtcars)
```


### tidycensus
```{r}
load_variables(2020, "acs5", cache = TRUE)

get_acs(geography="county",state="PA",year=2020,variables = c("B01002_001"),geometry = TRUE)

median_age_2020<-get_acs(geography="county",state="PA",year=2020,variables = c("B01002_001"),geometry = TRUE)

mapview(median_age_2020, zcol="estimate")

median_age_2020_tract<-get_acs(geography="tract",state="PA",year=2020,variables = c("B01002_001"),geometry = TRUE)

mapview(median_age_2020_tract, zcol="estimate")
```

